Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-18 arxiv/proof_examples.tex
section D wip

We follow a similar logic as in the proof of Theorem \ref{thm_many_tasks}.
We first characterize the global minimizer of $f(A, B)$ in the random-effects model.
Based on the characterization, we reduce the prediction loss of hard parameter sharing to the asymptotic limits provided in Theorem \ref{thm_main_RMT}.
Then, we prove Corollary \ref{cor_MTL_loss} based on the bias and variance limits.

@@

We assume that $A_1$ and $A_2$ are both nonzero.
Otherwise, we add a tiny amount of perturbation $\delta$ to them and the result remains the same.

@@

We denote $\hat \Sigma(x):= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}$.

@@

		 + & \left\| X^{(2)} \hat \Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (x\beta^{(1)}-x^2\beta^{(2)})+ \left(X^{(2)}\hat\Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2\times n_2}\right)\epsilon^{(2)} \right. \nonumber\\
		 &\left. + x X^{(2)}\hat \Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2.
		\label{eq_mtl_A12}

@@

The conditional expectation of $g(x)$ over $\epsilon^{(1)}$ and $\epsilon^{(2)}$ is
%\HZ{could you align the Eq so that they look better on page?}
\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta^{(1)},\beta^{(2)}}
&= \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})\right\|^2 \nonumber\\
&+x^2 \left\| X^{(2)} \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta^{(1)}-x\beta^{(2)})\right\|^2  \nonumber\\
& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2\epsilon^{(1)}} \nonumber\\
&+\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\hat \Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2\times n_2}\right)^2\epsilon^{(2)}}   \nonumber\\
&+  x^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)}\hat \Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)}} \nonumber\\
& +x^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)}\hat \Sigma(x)^{-1}  (X^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)}  }. \label{Eg(x)}
Using the following identity
(X^{(1)})^\top X^{(1)} \hat \Sigma(x)^{-1}(X^{(2)})^\top X^{(2)} & =\left( x^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \\
&= (X^{(2)})^\top X^{(2)} \hat \Sigma(x)^{-1}(X^{(1)})^\top X^{(1)},
we can simplify that
 & \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})\right\|^2  +x^2 \left\| X^{(2)} \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta^{(1)}-x\beta^{(2)})\right\|^2 \\
=&(\beta^{(1)}-x\beta^{(2)})^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1} (X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)}  (\beta^{(1)}-x\beta^{(2)})\\
&+(\beta^{(1)}-x\beta^{(2)})^\top x^2 (X^{(1)})^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)} )^\top X^{(2)}  \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta^{(1)}-x\beta^{(2)})\\
%&=(\beta^{(1)}-x\beta^{(2)})^\top\left[ x^2 (X^{(1)})^\top X^{(1)}+(X^{(2)})^\top X^{(2)}  \right]\hat\Sigma(x)^{-1} (X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})\\
=&(\beta^{(1)}-x\beta^{(2)})^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)}).
Using this identity, we can simplify equation \eqref{Eg(x)}  to

@@

		=& (\beta^{(1)}-x\beta^{(2)})^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})  \\
		&+ \sigma^2\tr\left[ \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2+ \left(X^{(2)}\hat\Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2\times n_2}\right)^2 \right]\\
&+ 2 x^2 \sigma^2 \tr\left[ \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\

@@

For the random-effects model, recall that the entries of $(\beta^{(1)}- x\beta^{(2)})\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(x-1)^2\kappa^2 +(1+x^2) d^2]$.
Hence, 

@@

&\exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta^{(1)},\beta^{(2)}}{g(x) \mid X_1, X_2} \nonumber\\

@@

&=\frac{(x-1)^2\kappa^2 + (x^2+1)d^2}{p} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma(x)^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}

@@

\paragraph{Part $1$: characterizing the global minimum of $f(A, B)$.}
Let $\hat{x}$ denote the minimizer of $g(x)$.
We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.
This gives us the global minimum of $f(A, B)$, combined with the local optimality condition for $\hat{B}$.

First, we show that $g(x)$ and $\ex{g(x) \mid X^{(1)}, X^{(2)}}$ are close using standard concentration bounds.
\begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability
$$\left| g(x) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta^{(1)},\beta^{(2)}}{g(x) \mid X_1, X_2 }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right)$$
for any small constant $c>0$.
 \end{claim}
 \begin{proof}
By Fact \ref{fact_minv} (ii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
 \ee
 and
 \be\label{op_Sig1}
\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
 \ee
where we used that $3\le n_1/p\le \tau^{-1}$ and $3\le n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

 Now we expand the first term on the right-hand side of \eqref{eq_mtl_A12}:
 \begin{align}
 & \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)})+ \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)} \right. \nonumber\\
		& \left. + x X^{(1)}\hat \Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2=h_1(x) + h_2(x) +h_3(x) + 2h_4(x) + 2h_5(x) + 2h_6(x) ,\label{expand_6}
 \end{align}
 where
 \begin{align*}
&h_1(x):=  (\beta^{(1)}-x\beta^{(2)})^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})  ,\\
& h_2(x) := {(\epsilon^{(1)})^\top \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2\epsilon^{(1)}}  ,\\
& h_3(x):=  x^2 { (\epsilon^{(2)} )^\top X^{(2)} \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)}} , \\
& h_4(x):=  (\epsilon^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}), \\
&h_5(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}),\\
&h_6(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)}.
\end{align*}
Next we estimate each term using Lemma \ref{largedeviation} in Appendix \ref{app_tool}.

For $h_1(x)$, using Lemma \ref{largedeviation} in Appendix \ref{app_tool} and the fact that the entries of $(\beta^{(1)}-x\beta^{(2)})\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(x-1)^2\kappa^2 + (x^2+1)d^2]$, we obtain the following estimate with high probability for any small constant $c>0$:
\begin{align}
&\left|h_1(x) - \exarg{\beta^{(1)},\beta^{(2)}}{h_1(x) \mid X_1, X_2}\right|\nonumber \\
&\le p^{c}\cdot p^{-1}[(x-1)^2\kappa^2 + (x^2+1)d^2]\cdot \left\| (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\|_F \nonumber\\
&\le p^{c}\cdot p^{-1}[(x-1)^2\kappa^2 + (x^2+1)d^2]\cdot p^{1/2} \left\| (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\| \nonumber \\
&\lesssim p^{1/2+c}\cdot \frac{(x-1)^2\kappa^2 + (x^2+1)d^2}{(x^2+1)^2}\lesssim p^{1/2+c}(\kappa^2 + d^2).\label{eq_hA1}
\end{align}
 Here in the third step we used \eqref{op_X12} and \eqref{op_Sig1} to bound the operator norm:
\begin{align*}
 \left\| (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\| & \le \left\| (X^{(2)})^\top X^{(2)}\right\|^2 \cdot \left\| (X^{(1)})^\top X^{(1)}\right\|\cdot \left\| \hat\Sigma(x)^{-1} \right\|^2 \\
 & \lesssim \frac{p}{(x^2+1)^2}.
\end{align*}
Similarly, using Lemma \ref{largedeviation}, the fact that the entries of $\epsilon^{(1)},\epsilon^{(2)}\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variances $\sigma^2$, and the bounds \eqref{op_X12}-\eqref{op_Sig1}, we can obtain that with high probability,
\be\label{eq_hA23}\left|h_2(x) - \exarg{\epsilon^{(1)}}{h_2(x) \mid X_1, X_2}\right| \lesssim p^{1/2+c}\sigma^2, \quad \left|h_3(x) - \exarg{\epsilon^{(2)}}{h_3(x) \mid X_1, X_2}\right| \lesssim p^{1/2+c}\sigma^2.\ee
For $h_4(x)$, using Lemma \ref{largedeviation} in Appendix \ref{app_tool}, we obtain the following estimate with high probability for any small constant $ c>0$:
\begin{align}
&\left|h_4(x)\right| \nonumber\\
&\le p^{c}\cdot \sigma \sqrt{p^{-1}[(x-1)^2\kappa^2 + (x^2+1)d^2]}\cdot \left\| \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\|_F\nonumber \\
&\le p^{c}\cdot \sigma \sqrt{p^{-1}[(x-1)^2\kappa^2 + (x^2+1)d^2]} \cdot p^{1/2} \left\| \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\|  \nonumber\\
&\le  p^{c}\cdot \sigma \sqrt{ (x-1)^2\kappa^2 + (x^2+1)d^2} \cdot \left\|  \id_{n_1\times n_1}-x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top\right\|  \left\|X^{(1)}\right\| \left\|\hat\Sigma(x)^{-1}\right\|\left\| (X^{(2)})^\top X^{(2)} \right\|  \nonumber\\
&\lesssim p^{1/2+c}\frac{\sigma \sqrt{ (x-1)^2\kappa^2 + (x^2+1)d^2}}{x^2+1} \lesssim p^{1/2+c}(\sigma^2 + \kappa^2 + d^2).\label{eq_hA4}
\end{align}
Above, in the fourth step we used $\left\|  \id_{n_1}-x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top\right\|\le 1$ and equations \eqref{op_X12} and \eqref{op_Sig1}. In the last step we used AM-GM inequality. With the same argument, we can show that
\be\label{eq_hA56} \left|h_5(x)\right| \le p^{1/2+c}(\sigma^2 + \kappa^2 + d^2), \quad \left|h_6(x)\right| \le p^{1/2+c}\sigma^2,\ee
with high probability for any small constant $c>0$. %Combining \eqref{eq_hA1}, \eqref{eq_hA23}, \eqref{eq_hA4} and \eqref{eq_hA56}, we conclude that

Finally, we can obtain similar estimates for the second term on the right-hand side of \eqref{eq_mtl_A12} as in  \eqref{eq_hA1}, \eqref{eq_hA23}, \eqref{eq_hA4} and \eqref{eq_hA56}. Hence the proof is complete.
 \end{proof}

Next, we show that $\hat x$ is close to 1.
\begin{claim}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability,
%There exists a constant $C>0$ such that
	\be\label{hatw_add1}|\hat x -1|\le  \frac{2d^2}{\kappa^2} + p^{-1/4+c}
	\ee
	for any small constant $c>0$.
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{claim}

>>>>

We follow a similar logic to the proof of Theorem \ref{thm_many_tasks}.
We first characterize the global minimizer of $f(A, B)$ in the random-effect model.
Based on the characterization, we reduce the prediction loss of hard parameter sharing to the bias-variance asymptotic limits.
Finally, we prove Corollary \ref{cor_MTL_loss} based on these limiting estimates.

@@

Without loss of generality, we assume that $A_1$ and $A_2$ are both nonzero.
Otherwise, the problem reduces to STL. % we add a tiny amount of perturbation $\delta$ to them and the result remains the same.

@@

We denote $\hat \Sigma(x)= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}$.

@@

		  & + \left\| X^{(2)} \hat \Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (x\beta^{(1)}-x^2\beta^{(2)})+ \left(X^{(2)}\hat\Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2\times n_2}\right)\epsilon^{(2)} \right. \nonumber\\
		  &\left. + x X^{(2)}\hat \Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2. \label{eq_mtl_A12}

@@

We have that the conditional expectation of $g(x)$ over $\epsilon^{(1)}$ and $\epsilon^{(2)}$ is

@@

%		=& (\beta^{(1)}-x\beta^{(2)})^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})  \\
%		&+ \sigma^2\tr\left[ \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2+ \left(X^{(2)}\hat\Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2\times n_2}\right)^2 \right]\\
%&+ 2 x^2 \sigma^2 \tr\left[ \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\

@@

The calculation is tedious but rather straightforward, so we leave the details to the reader.
In the random-effect model, recall that the entries of $(\beta^{(1)}- x\beta^{(2)})\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(x-1)^2\kappa^2 +(1+x^2) d^2]$.
Hence, by further taking expectation over $\beta^{(1)}$ and $\beta^{(2)}$, we obtain

@@

	&\ex{g(x) \mid X_1, X_2} \nonumber\\

@@

	=& ((x-1)^2\kappa^2 + (x^2+1)d^2){p^{-1}} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma(x)^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}

@@

%\HZ{could you align the Eq so that they look better on page?}
%\begin{align}
%\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta^{(1)},\beta^{(2)}}
%&= \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})\right\|^2 \nonumber\\
%&+x^2 \left\| X^{(2)} \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta^{(1)}-x\beta^{(2)})\right\|^2  \nonumber\\
%& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2\epsilon^{(1)}} \nonumber\\
%&+\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\hat \Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2\times n_2}\right)^2\epsilon^{(2)}}   \nonumber\\
%&+  x^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)}\hat \Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)}} \nonumber\\
%& +x^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)}\hat \Sigma(x)^{-1}  (X^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)}  }. \label{Eg(x)}
%\end{align}
%Using the following identity
%\begin{align*}
%(X^{(1)})^\top X^{(1)} \hat \Sigma(x)^{-1}(X^{(2)})^\top X^{(2)} & =\left( x^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \\
%&= (X^{(2)})^\top X^{(2)} \hat \Sigma(x)^{-1}(X^{(1)})^\top X^{(1)},
%\end{align*}
%we can simplify that
%\begin{align*}
% & \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})\right\|^2  +x^2 \left\| X^{(2)} \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta^{(1)}-x\beta^{(2)})\right\|^2 \\
%=&(\beta^{(1)}-x\beta^{(2)})^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1} (X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)}  (\beta^{(1)}-x\beta^{(2)})\\
%&+(\beta^{(1)}-x\beta^{(2)})^\top x^2 (X^{(1)})^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)} )^\top X^{(2)}  \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta^{(1)}-x\beta^{(2)})\\
%&=(\beta^{(1)}-x\beta^{(2)})^\top\left[ x^2 (X^{(1)})^\top X^{(1)}+(X^{(2)})^\top X^{(2)}  \right]\hat\Sigma(x)^{-1} (X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})\\
%=&(\beta^{(1)}-x\beta^{(2)})^\top(X^{(1)} )^\top X^{(1)}  \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)}).
%\end{align*}
%Using this identity, we can simplify equation \eqref{Eg(x)}  to

\paragraph{Part $1$: characterizing the global minimum of $f(A, B)$.}
Let $\hat{x}$ denote the global minimizer of $g(x)$.
We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.
This gives us the global minimum of $f(A, B)$, since $\hat{B}$ is given by $\hat{x}$ using local optimality conditions.
First, we show that $g(x)$ and $\ex{g(x) \mid X^{(1)}, X^{(2)}}$ are close using standard concentration bounds.
\begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, for any $x = O(1)$, we have that with high probability
$$\left| g(x) - \ex{g(x) \mid X^{(1)}, X^{(2)} }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right).$$
\end{claim}
\begin{proof}
 There are two terms in $g(A)$ from equation \eqref{eq_mtl_A12}.
 We will focus on dealing with the concentration error of the first term. The second term is similar to the first and we omit the details.
 For the first term, we expand into several equations under various situations involving the random noise and the random-effect model.
 \begin{align}
 & \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)})+ \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)} \right. \nonumber\\
		& \left. + x X^{(1)}\hat \Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2=h_1(x) + h_2(x) +h_3(x) + 2h_4(x) + 2h_5(x) + 2h_6(x), \label{expand_6}
 \end{align}
 where
 \begin{align*}
&h_1(x):=  (\beta^{(1)}-x\beta^{(2)})^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})  ,\\
& h_2(x) := {(\epsilon^{(1)})^\top \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2\epsilon^{(1)}}  ,\\
& h_3(x):=  x^2 { (\epsilon^{(2)} )^\top X^{(2)} \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)}} , \\
& h_4(x):=  (\epsilon^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}), \\
&h_5(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}),\\
&h_6(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)}.
\end{align*}
Next, we estimate each term using Lemma \ref{largedeviation} for random variables with bounded moment up to any order.
We first state several facts that will be commonly used in the proof.
By Fact \ref{fact_minv} (ii), we have that w.h.p. the operator norm of $X^{(1)}$ and $X^{(2)}$ are both bounded by $O(\sqrt{n})$.
Furthermore, the operator norm of $\hat\Sigma(x)^{-1}$ is bounded by $(x^2 + 1)^{-1} O(n_1 + n_2) = (x^2 + 1) O(p)$.
 % \be\label{op_X12}
%\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
% \ee
% and
% \be\label{op_Sig1}
%\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
% \ee
%where we used that $3\le n_1/p\le \tau^{-1}$ and $3\le n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

For $h_1(x)$, using Lemma \ref{largedeviation} and the fact that the entries of $(\beta^{(1)}-x\beta^{(2)})\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $b = p^{-1}((x-1)^2\kappa^2 + (x^2+1)d^2)$, we obtain the following estimate w.h.p.
\begin{align}
	&\left|h_1(x) - \exarg{\beta^{(1)},\beta^{(2)}}{h_1(x) \mid X_1, X_2}\right|\nonumber \\
\le& p^{c}\cdot p^{-1} b \cdot \left\| (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\|_F \nonumber\\
	\le& p^{c}\cdot p^{-1} b \cdot p^{1/2} \left\| (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\| \nonumber \\
	\lesssim& p^{1/2+c}\cdot \frac{b}{(x^2+1)^2}\lesssim p^{1/2+c}(\kappa^2 + d^2).\label{eq_hA1}
\end{align}
In the third step we use the operator norm bound of $X^{(1)}$, $X^{(2)}$, and $\hat{\Sigma}(x)^{-1}$. % as follows
%\begin{align*}
%	\left\| (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\| & \le \left\| (X^{(2)})^\top X^{(2)}\right\|^2 \cdot \left\| (X^{(1)})^\top X^{(1)}\right\|\cdot \left\| \hat\Sigma(x)^{-1} \right\|^2 \\
%  & \lesssim \frac{p}{(x^2+1)^2}.
%\end{align*}

For $h_2(x)$ and $h_3(x)$, since the entries of $\epsilon^{(1)},\epsilon^{(2)}\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variances $\sigma^2$, using Lemma \ref{largedeviation}, we obtain w.h.p.
\be\label{eq_hA23}\left|h_2(x) - \exarg{\epsilon^{(1)}}{h_2(x) \mid X_1, X_2}\right| \lesssim p^{1/2+c}\sigma^2, \quad \left|h_3(x) - \exarg{\epsilon^{(2)}}{h_3(x) \mid X_1, X_2}\right| \lesssim p^{1/2+c}\sigma^2.\ee
For $h_4(x)$, using Lemma \ref{largedeviation}, we obtain w.h.p.:
\begin{align}
			&\left|h_4(x)\right| \nonumber\\
	\le& p^{c}\cdot \sigma \cdot \sqrt{b / p}\left\| \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\|_F\nonumber \\
	\le& p^{c}\cdot \sigma \sqrt{b / p} \cdot p^{1/2} \left\| \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} \right\|  \nonumber\\
	\le&  p^{c}\cdot \sigma \sqrt{b} \cdot \left\|  x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top - \id_{n_1\times n_1} \right\| \cdot \left\|X^{(1)}\right\| \cdot \left\|\hat\Sigma(x)^{-1}\right\| \cdot \left\| (X^{(2)})^\top X^{(2)} \right\|  \nonumber\\
	\lesssim& p^{1/2+c}\frac{\sigma \sqrt{b}}{x^2+1} \lesssim p^{1/2+c}(\sigma^2 + \kappa^2 + d^2).\label{eq_hA4}
\end{align}
Above, in the fourth step we use the operator norm of $x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top - \id$ being at most one and the operator norm bound of $X^{(1)}$, $X^{(2)}$, and $\hat{\Sigma}(x)^{-1}$.
In the last step we use AM-GM inequality. Using the same argument, we can show that
$\left|h_5(x)\right| \le p^{1/2+c}(\sigma^2 + \kappa^2 + d^2)$ and $\left|h_6(x)\right| \le p^{1/2+c}\sigma^2$.
Combining the concentration error bound for $h_1(x), h_2(x), \dots, h_6(x)$, we complete the proof.
The second term of $g(A)$ can be dealt in similar ways and we omit the details.
\end{proof}
